Exhaustive dynamic analysis of reachability and stable states of biological regulatory networks modeled in Process Hitting using Answer Set Programming

### 2015-04-29 parts/comparaison.tex
modif de la version de pint

\subsection{Evaluation}
To assess the efficiency of our new approach,
we position ourselves with respect to existing methods dealing with different biological models.
We have chosen the following tools, that are detailed below: 
\textsc{GINsim}\footnote{\textsc{GINsim} version 2.4 alpha: \url{http://ginsim.org/}} (Gene Interaction Network Simulation)~\cite{gonzalez2006ginsim,naldi2009logical,naldi2007decision},
\textsc{LibDDD}\footnote{\textsc{LibDDD} version 1.8: \url{http://move.lip6.fr/software/DDD/}}
(Library of Data Decision Diagrams)~\cite{thierry2009hierarchical,colange2013towards},
\textsc{Pint}\footnote{\textsc{Pint} version 2015-02-11: \url{http://loicpauleve.name/pint/}}~\cite{PMR12-MSCS}
and the method for CTL model-checking proposed by Rocca \textit{et al.} in~\cite{roccaasp},
which was developed also in ASP but for states transitions networks.
Each method uses a specific kind of representation\footnote{When available, we used the converters included into \textsc{Pint} for these translations.}:
Thomas models (a particular kind of logical regulatory networks) for \textsc{GINsim},
instantiable transition systems for \textsc{LibDDD},
state transition networks for the method of Rocca \textit{et al.}
and Process Hitting (PH) for \textsc{Pint} as well as for our method.

@@

In the previous sections,
we developed new methods in ASP to check dynamical properties,
namely identifying stable states and finding all possible paths to reach a given goal.
Compared to some other methods describes above
%and the method \textsc{ASP-Thomas}
(\textsc{GINsim} and \textsc{LibDDD}) our method is relatively faster and also permits to study larger networks
(up to $2^{73}$ states in our tests). We exclusively study networks modeled in Process Hitting. This new formalism for network modeling is a restriction of synchronous automata and thus allows to represent any kind of dynamical model. Moreover it is easy to implement the dynamics of the PH into ASP.
However, our ASP program may still be non-conclusive
in the cases where the given goal is not
reachable and the model contains loops in its dynamics
(which happens in almost every biological model).
In this case, the program will compute all infinite paths in these loops,
and never reach a goal or a fixed point.
It is still possible, however, to limit the number of iterations to an arbitrary
maximum which will be eventually reached in the case of an endless loop.
This is possible with the option ``\texttt{-{}-imax=n}'' of \textsc{iClingo},
where \texttt{n} is the maximum number of steps.
%%
%\begin{lstlisting}[numbers=none]
%iclingo find\_path.lp <network\_name>.lp -- imax=n
%\end{lstlisting}
%%

>>>>

\subsection{Evaluation}
To assess the efficiency of our new approach,
we position ourselves with respect to existing methods dealing with different biological models.
We have chosen the following tools, that are detailed below: 
\textsc{GINsim}\footnote{\textsc{GINsim} version 2.4 alpha: \url{http://ginsim.org/}} (Gene Interaction Network Simulation)~\cite{gonzalez2006ginsim,naldi2009logical,naldi2007decision},
\textsc{LibDDD}\footnote{\textsc{LibDDD} version 1.8: \url{http://move.lip6.fr/software/DDD/}}
(Library of Data Decision Diagrams)~\cite{thierry2009hierarchical,colange2013towards},
\textsc{Pint}\footnote{\textsc{Pint} version 2015-11-14: \url{http://loicpauleve.name/pint/}}~\cite{PMR12-MSCS}
and the method for CTL model-checking proposed by Rocca \textit{et al.} in~\cite{roccaasp},
which was developed also in ASP but for states transitions networks.
Each method uses a specific kind of representation\footnote{When available, we used the converters included into \textsc{Pint} for these translations.}:
Thomas models (a particular kind of logical regulatory networks) for \textsc{GINsim},
instantiable transition systems for \textsc{LibDDD},
state transition networks for the method of Rocca \textit{et al.}
and Process Hitting (PH) for \textsc{Pint} as well as for our method.

@@

In the previous sections,
we developed new methods in ASP to check dynamical properties,
namely identifying stable states and finding all possible paths to reach a given goal.
Compared to some other methods describes above
%and the method \textsc{ASP-Thomas}
(\textsc{GINsim} and \textsc{LibDDD}) our method is relatively faster and also permits to study larger networks
(up to $2^{73}$ states in our tests). We exclusively study networks modeled in Process Hitting. This new formalism for network modeling is a restriction of synchronous automata and thus allows to represent any kind of dynamical model. Moreover it is easy to implement the dynamics of the PH into ASP.
However, our ASP program may still be non-conclusive
in the cases where the given goal is not
reachable and the model contains loops in its dynamics.
In this case, the program will compute all infinite paths in these loops,
and never reach a goal or a fixed point.
It is still possible, however, to limit the number of iterations to an arbitrary
maximum which will be eventually reached in the case of an endless loop.
This is possible with the option "\texttt{-{}-imax=n}'' of \textsc{iClingo},
where \texttt{n} is the maximum number of steps.
%%
%\begin{lstlisting}[numbers=none]
%iclingo find\_path.lp <network\_name>.lp -- imax=n
%\end{lstlisting}
%%
